\chapter*{Preface}
\setheader{Preface}
\addcontentsline{toc}{chapter}{Preface}

You are currently reading the thesis ``\makeatletter\@title\makeatother''. It has been written in order to obtain the degree of Bachelor of Science in Applied Mathematics and Applied Physics at the Delft University of Technology. I engaged in researching and writing this thesis from the 24th of February to the 1st of July in the year 2022.

From a young age I was captivated by prime numbers. At first, I studied them on paper, but through lack of brain power I started to write computer code. As I got older the efficiency of my computer code increased and from it originated my fascination for programming and optimizing algorithms.
Researching and enhancing the calculation of the Rayleigh integral using my knowledge obtained inside and outside this Bachelor degree was therefore an interesting quest and a joyful challenge.

I would like to thank my supervisors, Kees Vuik and Eric Verschuur, as they were of great help during my bachelor thesis. I was especially thankful for the freedom they gave me to explore the topics I found appealing, whilst making sure that the topics remained relevant for present-day applications.
I am also grateful for Yoeri Dijkstra and Koen van Dongen, the other members of my thesis committee, for attending the defense of my thesis and assessing this report.
In addition, Leo Hoogerbrugge, currently Eric's PhD-student, provided great knowledge and assistance in our weekly meetings.

Further gratitude goes out towards my parents, my brothers, and friends. The feedback I received about the content, code and layout of this thesis was very helpful and more than appreciated.

\begin{flushright}
{\makeatletter\itshape
    \@author \\
    The Hague, July 2022
\makeatother}
\end{flushright}

\blfootnote{The figure at the cover is Figure \ref{fig:real_3d_projected_wavefield}, displaying an artificial integrand of the Rayleigh integral. In this thesis we develop various methods to integrate this function, which we then compare at the end.}
